---
title: "Lab 03"
subtitle: "Inference for linear regression"
date: "September 28, 2026"
date-format: long
footer: "[🔗 sta221-fa26.github.io](https://sta221-fa26.github.io)"
logo: "../images/logo.png"
format: 
  revealjs: 
    theme: slides.scss
    multiplex: false
    transition: fade
    slide-number: true
    incremental: false 
    chalkboard: true
    include-before: [ '<script type="text/x-mathjax-config">MathJax.Hub.Config({tex2jax: {enableAssistiveMml: false}});</script>']
html-math-method:
  method: mathjax
  url: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-mml-chtml.js"

execute:
  freeze: auto
  echo: true
knitr:
  opts_chunk: 
    R.options:      
    width: 200
---

# Welcome

## Goals

- Lab 03: Inference for linear regression
- Project proposal

```{r}
#| include: false

library(tidyverse)
library(tidymodels)
library(knitr)

football <- read_csv("data/ncaa-football-exp.csv")
```

## Hypothesis testing example {.midi}

```{r}
#| echo: false

lm(total_exp_m ~ enrollment_th + type, data = football) |>
  tidy() |>
  kable(digits = 3)
```

:::: question
::: midi
We wish to test whether there is evidence of a linear relationship between enrollment and football expenditure after accounting for the type of institution.

1.  State the null and alternative hypotheses in words and mathematical notation.
2.  State how the test statistic is computed from the other values in the output. Interpret the test statistic in the context of the data.
3.  State the distribution used to compute the p-value and interpret it in the context of the data. \*Note: There are `r nrow(football)` observations in the data.\*
4.  State your conclusion in the context of the data.
:::
::::

## Lab 03: Inference for linear regression

Today’s lab focuses on using statistical inference to draw conclusions about factors associated with weekly avocado sales.

<br>

🔗 [sta221-fa26.github.io/labs/lab-03.html](../labs/lab-03.html)

<br>

If you finish the lab early, use the remaining time to work on the project proposal.
