---
title: "Variance Inflation Factors"
format: html
bibliography: references.bib
---

```{r}
#| echo: true
#| message: false

# load packages
library(tidyverse)  
library(tidymodels)  
library(knitr)       
library(rms)

rail_trail <- read_csv("data/rail_trail.csv")
```

Here we explain the connection between the Variance Inflation Factor (VIF) and $\mathbf{C} = (\mathbf{X}^\mathsf{T}\mathbf{X})^{-1}$. This explanation is motivated by Chapter 3 of @montgomery2021introduction.

## Unit length scaling

We have talked about standardizing predictors, such that

$$x_{ij_{std}} = \frac{x_{ij} - \bar{x_j}}{s_{x_{j}}}$$ such that $\bar{x}_j$ is the mean and $s_{x_{j}}$ is the standard deviation of the predictor $x_j$.

The standardized predictors have a mean of 0 and variance of 1.

Another common type of scaling is **unit length scaling**. We will denote these scaled predictors as $w_j$. We apply this scaling on both the predictor and response variable in the following way.

$$w_{ij} = \frac{x_{ij} - \bar{x_j}}{\sqrt{s_{jj}}}$$ where $s_{jj} = \sum_{i = 1}^{n}(x_{ij} - \bar{x}_j)^2$

The scaled response variable, denoted $y^0$ is

$$y_i^0 = \frac{y_i - \bar{y}}{\sqrt{SST}}$$

where $SST$ is the sum of squares total, $\sum_{i=1}^n(y_i - \bar{y})^2$

We will use the scaled predictor and response variable to show the relationship between $\mathbf{C} = (\mathbf{X}^\mathsf{T}\mathbf{X})^{-1}$ and the formula for VIF. More specifically, that

$$
C_{jj} = VIF_{j} = \frac{1}{1 - R^2_j}
$$

when we use the scaled predictors and response.

## Variance inflation factor

We will use the `rail_trail` data from the notes to illustrate this connection. We will focus on the predictors `hightemp`, `avgtemp`, and `precip`.

We begin by creating unit length scaled versions of the variables.

```{r}

hightemp_norm <- sum((rail_trail$hightemp - mean(rail_trail$hightemp))^2)
avgtemp_norm <- sum((rail_trail$avgtemp - mean(rail_trail$avgtemp))^2)
precip_norm <- sum((rail_trail$precip - mean(rail_trail$precip))^2)
volume_norm <- sum((rail_trail$volume - mean(rail_trail$volume))^2)

rail_trail <- rail_trail |>
  mutate(hightemp_scaled = (hightemp - mean(hightemp)) / hightemp_norm^.5,
         avgtemp_scaled = (avgtemp - mean(avgtemp)) / avgtemp_norm^.5, 
         precip_scaled = (precip - mean(precip)) / precip_norm^.5,
         volume_scaled = (volume - mean(volume)) / volume_norm^.5
         )
```

The matrix $\mathbf{W}^\mathsf{T}\mathbf{W}$ is equivalent to the correlation matrix for these predictors.

```{r}
# use -1 to remove the intercept for the correlation matrix
W <- model.matrix(volume_scaled ~ hightemp_scaled + avgtemp_scaled + precip_scaled - 1, data = rail_trail)

t(W)%*%W
```

When we fit a model using the unit length scaling for the response and predictor variables, we would expect $Var(\hat{\beta}_j) / \hat{\sigma}^2_{\epsilon} \approx 1$. As we see below, however, these values are greater than 1.

```{r}
trail_model_scaled <- lm(volume_scaled ~ hightemp_scaled + avgtemp_scaled + precip_scaled , data = rail_trail)

beta_se <- tidy(trail_model_scaled)$std.error
sigma <- glance(trail_model_scaled)$sigma

beta_se^2 / sigma^2
```

(You can ignore the first element, which represents the intercept).

These values show how much the standard errors of the coefficients are inflated given the correlation between the predictors (the off diagonal elements of $\mathbf{W}^\mathsf{T}\mathbf{W}$). The amount by which the standard errors are inflated are called the **variance inflation factors (VIF)**.

Under this model using the unit-length-scaled predictors and response, we see these **variance inflation factors** are equal to the diagonal elements of $\mathbf{C} = (\mathbf{W}^\mathsf{T}\mathbf{W})^{-1}$.

```{r}
C <- solve(t(W) %*% W)
diag(C)
```

```{r}
# compute 
vif(lm(volume ~ hightemp + avgtemp + precip, data = rail_trail))
```

Thus,

$$C_{jj} = VIF_{j} = \frac{1}{1 - R^2_j}$$

## References
