---
title: "AE 01: Model evaluation"
subtitle: "Income inequality, healthcare expenditure, and life expectancy"
date: "September 1, 2026"
editor: visual
toc: true
execute: 
  message: false
  warning: false
bibliography: references.bib
---

::: callout-important
Go to the [course GitHub organization](https://github.com/sta221-fa26) and locate your `ae-01` repo to get started.

Render, commit, and push your responses to GitHub. This is good practice for the workflow you will use in assignments.

The AEs are for practice only. They are not graded.
:::

```{r}
#| label: load-packages
#| message: false

library(tidyverse)
library(tidymodels) 
library(knitr) 
```

# Data

The data set comes from @zarulli2021 who analyze the effects of a country's healthcare expenditure and other factors on the country's life expectancy. The data are originally from the [Human Development Database](http://hdr.undp.org/en/data) and [World Health Organization](https://apps.who.int/nha/database/).

- `life_exp`: The average number of years that a newborn could expect to live, if he or she were to pass through life exposed to the sex- and age-specific death rates prevailing at the time of his or her birth, for a specific year, in a given country, territory, or geographic income_inequality. ( from the [World Health Organization](https://www.who.int/data/gho/indicator-metadata-registry/imr-details/65#:~:text=Definition%3A,%2C%20territory%2C%20or%20geographic%20area.))

- `income_inequality`: Measure of the deviation of the distribution of income among individuals or households within a country from a perfectly equal distribution. A value of 0 represents absolute equality, a value of 100 absolute inequality (based on Gini coefficient). (from @zarulli2021)

- `health_expend`: Per capita current spending on on healthcare good sand services, expressed in respective currency - international Purchasing Power Parity (PPP) dollar ([from the World Health Organization](https://www.who.int/data/gho/indicator-metadata-registry/imr-details/4952))

- `health_pct_gdp`: Spending on healthcare goods and services, expressed as a percentage of GDP. It excludes capital health expenditures such as buildings, machinery, information technology and stocks of vaccines for emergency or outbreaks (from @zarulli2021).

```{r}
#| label: load-data

life_exp_data <- read_csv("data/life_exp.csv")

```

# Part 1

## Exercise 1

Fit a model using income equality to understand variability in life expectancy. Neatly display the results using 3 decimal places.

```{r}
#| label: ex-1

# add code here

```

## Exercise 2

- Interpret the slope in the context of the data.

- Does the intercept have a meaningful interpretation? If so, interpret it in the context of the data. Otherwise, explain why not.

# Part 2

We now want to understand the relationship between a country's healthcare expenditure and its life expectancy. The data set contains two measures for healthcare expenditure: `health_expend` and `health_pct_gdp`.

## Exercise 3

Fit a model using `health_expend` to understand variability in `life_exp`. Compute $R^2$ and $RMSE$ for this model.

```{r}
#| label: ex-3

# add code here

```

- Interpret $R^2$ in the context of the data.

- Interpret $RMSE$ in the context for the data.

## Exercise 4

Which measure of healthcare expenditure would you choose as a predictor of life expectancy - `health_expend` or `health_pct_gdp`? Briefly explain, using $R^2$ and/or $RMSE$ to support your choice.

```{r}
#| label: ex-4

# add code here

```

# Wrapping up

::: callout-important
Once you've completed the AE:

- Render the document to produce the PDF with all of your work from today's class.
- Push all your work to your AE repo on GitHub. You're done! 🎉
:::
