---
title: "AE 04: Multicollinearity"
date: "February 24, 2026"
editor: visual
---

::: callout-important
Go to the [course GitHub organization](https://github.com/sta221-sp26) and locate your `ae-04` repo to get started.
:::

```{r}
#| label: load-packages
#| warning: false
#| message: false

library(tidyverse)
library(knitr)
library(tidymodels)
library(rms) #calculate VIF
```

# Introduction

The Pioneer Valley Planning Commission (PVPC) collected data at the beginning a trail in Florence, MA for ninety days from April 5, 2005 to November 15, 2005. Data collectors set up a laser sensor, with breaks in the laser beam recording when a rail-trail user passed the data collection station. The data were collected from by [Pioneer Valley Planning Commission](http://www.fvgreenway.org/pdfs/Northampton-Bikepath-Volume-Counts%20_05_LTA.pdf) via the **mosaicData** package.

We will use the following variables in this analysis:

**Response**

-   `volume` estimated number of trail users that day (number of breaks recorded)

**Predictors**

-   `hightemp` daily high temperature (in degrees Fahrenheit)

-   `avgtemp` average of daily low and daily high temperature (in degrees Fahrenheit)

-   `season` one of “Fall”, “Spring”, or “Summer”

-   `precip` measure of precipitation (in inches)

```{r}
#| label: load-data
#| message: false
#| eval: false

rail_trail <- read_csv("data/rail_trail.csv")
```

# Part 1

## Exercise 1

-   Fit the regression model using high temperature, average temperature, season, and precipitation to predict volume.

-   Are there any coefficients that may be not what you expected?

## Exercise 2

Use the formula

$$
VIF_j = \frac{1}{1 - R^2_j} 
$$

to calculate the VIF for `avgtemp`.

## Exercise 3

Based on the VIF from the previous exercise, does `avgtemp` have a linear dependency with one or more other predictors? Explain.

## Exercise 4

-   Use the `vif` function to compute VIF for all the predictors in Exercise 1.

-   Are there predictors with near-linear dependencies? If so, which ones?

# Part 2

## Exercise 5

Let's address the issue of multicollinearity. Choose a strategy to address the multicollinearity. Apply it, then use relevant statistics to select a final model.

# Wrapping up

::: callout-important
Once you've completed the AE:

-   Render the document to produce the PDF with all of your work from today's class.
-   Push all your work to your AE repo on GitHub. You're done! 🎉
:::
