---
title: "AE 02: Multiple linear regression"
subtitle: "Peer-to-peer lending"
date: "January 27. 2026"
editor: visual
---

::: callout-important
This AE will be completed through code chunks in the [lecture slides](../slides/06-predictor-types.html).
:::

# Packages

```{r}
#| label: load-packages
#| message: false

library(tidyverse)   
library(tidymodels)   
library(openintro)    
library(knitr)       
```

# Data

Today's data is a sample of 50 loans made through a peer-to-peer lending club. The data is in the `loan50` data frame in the **openintro** R package.

We will focus on the following variables:

-   `annual_income_th`: Annual income (in \$1000s)
-   `debt_to_income`: Debt-to-income ratio, i.e. the percentage of a borrower's total debt divided by their total income
-   `verified_income`: Whether borrower's income source and amount have been verified (`Not Verified`, `Source Verified`, `Verified`)
-   `interest_rate`: Interest rate for the loan

The goal of this analysis is to use the annual income, debt-to-income ratio, and income verification to understand variability in the interest rate on the loan.

We'll start with data prep to rescale annual income to \$1000's and recode `verified_income` to fix an issue with the underlying data.

```{r}
#| label: data-prep

loan50 <- loan50 |>
   mutate(annual_income_th = annual_income / 1000, 
          verified_income = 
            case_when(verified_income == "Not Verified" ~ "Not Verified",
                      verified_income == "Source Verified" ~ "Source Verified",
                      verified_income == "Verified" ~ "Verified"),
          verified_income = as_factor(verified_income)
   )                    
```

```{r}
#| label: glimpse-data

glimpse(loan50)
```

# Part 1

## Exercise 1

We'll start by fitting a model in which we include all levels of `verified_income`.

-   Fit a model using `debt_to_income`, `annual_income_th`, and the indicator variables created below to predict `interest_rate`.

-   What do you notice about the model output? Why did this happen?

```{r}
#| label: make-indicators

loan50 <- loan50 |>
  mutate(
    not_verified = factor(if_else(verified_income == "Not Verified", 1, 0)),
    source_verified = factor(if_else(verified_income == "Source Verified", 1, 0)),
    verified = factor(if_else(verified_income == "Verified", 1, 0))
  )
```

```{r}
#| label: fit-model-all-indicators 

# add code here

```

## Exercise 2

Now let's take a look at the design matrix for the model with predictors `debt_to_income`, `annual_income_th`, and `verified_income`.

How does R choose the baseline level by default?

```{r}
#| label: ex-1

## add code here

```

## Exercise 3

What is the intercept for individuals with

-   Not verified income?

-   Source verified income?

-   Verified income?

# Part 2

## Exercise 4

Fit the model with the predictors `debt_to_income`, `annual_income_th`, `verified_income` , and the interaction between `annual_income_th` and `verified_income`.

Neatly display the model results using 3 digits.

```{r}
#| label: ex-2

# add code here

```

## Exercise 5

-   Write the estimated regression equation for the people with `Not Verified` income.

-   Write the estimated regression equation for people with `Verified` income.

## Exercise 6

In general, how do

-   indicators for categorical predictors impact the model equation?

-   interaction terms impact the model equation?

# LaTex

Sometimes, you will need to include mathematical notation in your document. There are two ways you can display mathematics in your document:

**Inline:** Your mathematics will display within the line of text.

-   Use `$` to start and end your LaTex syntax. You can also use the menu: *Insert -\> LaTex Math -\> Inline Math*.

-   *Example*: The text `The simple linear regression model is $\mathbf{y} = \mathbf{X}\boldsymbol{\beta} + \boldsymbol{\epsilon}$` produces\

    The simple linear regression model is $\mathbf{y} = \mathbf{X}\boldsymbol{\beta} + \boldsymbol{\epsilon}$

\
**Displayed:** Your mathematics will display outside the line of text

-   Use a `$$` to start and end your LaTex syntax. You can also use the menu: *Insert -\> LaTex Math -\> Display Math*.

-   *Example*: The text `The estimated regression equation is $$\hat{\mathbf{y}} = \mathbf{X}\hat{\boldsymbol{\beta}}$$`produces\

    The estimated regression equation is

$$
\hat{\mathbf{y}} = \mathbf{X}\hat{\boldsymbol{\beta}}
$$

::: callout-tip
[Click here](https://oeis.org/wiki/List_of_LaTeX_mathematical_symbols) for a quick reference of LaTex code.
:::

# Submission

::: callout-important
To submit the AE:

-   Render the document to produce the PDF with all of your work from today's class.
-   Push all your work to your AE repo on GitHub. You're done! 🎉
:::
