---
title: "AE 02: Distribution of coefficients"
subtitle: "NCAA Football Expenditures"
date: "September 22, 2026"
editor: visual
---

::: callout-important
Go to the [course GitHub organization](https://github.com/sta221-fa26) and locate your `ae-02` repo to get started.
:::

# Set up

Run the code below to load the required packages and data.

```{r}
#| label: load-packages-data
#| message: false

library(tidyverse)
library(tidymodels)
library(knitr)

football <- read_csv("data/ncaa-football-exp.csv")
```

## Data: NCAA football expenditures

Today's data set comes from [Equity in Athletics Data Analysis](https://ope.ed.gov/athletics/#/datafile/list) and includes information about sports expenditures and revenues for colleges and universities in the United States. This data set was featured in a [March 2022 Tidy Tuesday](https://github.com/rfordatascience/tidytuesday/blob/master/data/2022/2022-03-29/readme.md).

We will focus on the 2019 - 2020 season expenditures on football for institutions in the NCAA - Division 1 FBS. The following variables are used in this analysis:

- `total_exp_m`: Total expenditures on football in the 2019 - 2020 academic year (in millions USD)

- `enrollment_th`: Total student enrollment in the 2019 - 2020 academic year (in thousands)

- `type`: institution type (Public or Private)

# Regression model

```{r}
exp_fit <- lm(total_exp_m ~ enrollment_th + type, data = football)

tidy(exp_fit)|> 
  kable(digits = 3)
```

# Part 1: Distribution of $\hat{\boldsymbol{\beta}}$

Our ultimate goal is to use statistical inference to draw conclusions about the relationship between student enrollment and spending on football programs, after accounting for the institution type.

## Exercise 1

We will use the vector of responses $\mathbf{y}$ and the design matrix $\mathbf{X}$ to compute the values needed for inference.

Obtain $\mathbf{y}$ and $\mathbf{X}$ from the football data frame. What are their dimensions?

## Exercise 2

Use $\mathbf{y}$ and $\mathbf{X}$ to compute the estimated coefficients $\hat{\boldsymbol{\beta}}$.

## Exercise 3

- Compute the residuals using the matrix/vector notation. You do not need to print the residuals.

- Use the residuals to compute $\hat{\sigma}^2$.

## Exercise 4

- Compute $Var(\hat{\boldsymbol{\beta}})$. What are its dimensions?

- Explain how the standard error for $\hat{\beta}_\text{enroll}$, the estimated coefficient of `enrollment_th`, is computed from $Var(\hat{\boldsymbol{\beta}})$.

## Exercise 5

The standard error of $\hat{\beta}_j$ is computed as $\sqrt{Var(\hat{\beta}_j)}$ . Interpret the standard error for $\hat{\beta}_\text{enroll}$ in the context of the data.

# Wrapping up

::: callout-important
Once you've completed the AE:

- Render the document to produce the PDF with all of your work from today's class.
- Push all your work to your AE repo on GitHub. You're done! 🎉
:::
