---
title: "Lab 04: Exam 01 Review"
html-math-method:
  method: mathjax
  url: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-mml-chtml.js"
editor: visual
execute:
  warning: false
  message: false
---

::: callout-important
# Due date

This lab is due at the end of your lab session on **October 5.** Be sure to complete the following by the end of your lab:

- Push final `.qmd` and `.pdf` files to your GitHub repo
- Submit final `.pdf` file on Gradescope

**No late work is accepted for labs.**
:::

# Getting Started

- Go to the [sta221-fa26](https://www.github.com/sta221-fa26) organization on GitHub. Click on the repo with the prefix **lab-04**. It contains the starter documents you need to complete the lab.

- Clone the repo and start a new project in RStudio. See the [Lab 00 instructions](../labs/lab-00.html) for details on cloning a repo and starting a new project in R.

- Each person on the team should clone the repository and open a new project in RStudio. Throughout the lab, each person should get a chance to make commits and push to the repo

# Packages

You will need the following packages for the lab:

```{r}
#| label: packages

library(tidyverse)
library(tidymodels) 
library(knitr)
library(patchwork)

# add other packages as needed
```

# Restaurant tips

**What factors are associated with the amount customers tip at a restaurant?** To answer this question, we will use data collected in 2011 by a student at St. Olaf who worked at a local restaurant.[^1]

[^1]: Dahlquist, Samantha, and Jin Dong. 2011. "The Effects of Credit Cards on Tipping." Project for Statistics 212-Statistics for the Sciences, St. Olaf College.

The variables we'll focus on for this analysis are

- `Tip`: amount of the tip
- `Party`: number of people in the party
- `Age`: Age of the payer (`Middle`, `Yadult`, `SenCit`)

View the data set to see the remaining variables.

```{r}
#| message: false

tips <- read_csv("data/tip-data.csv")
```

## Exploratory data analysis

```{r}
#| label: univariate-eda
#| warning: false

p1 <- ggplot(data = tips, aes(x = Tip)) + 
  geom_histogram(color = "white", binwidth = 2) +
  labs(x = "Tips ($)",
       title = "Tips at local restaurant")

p2 <- ggplot(data = tips, aes(x = Party)) + 
  geom_histogram(color = "white") +
  labs(x = "Party",
       title = "Number of diners in party") +
  xlim(c(0, 10))

p3 <- ggplot(data = tips, aes(x = Age)) + 
  geom_bar(color = "white") +
  labs(x = "",
       title = "Age of Payer") 

p1 / (p2 + p3)
```

```{r}
#| label: bivariate-eda

p4 <- ggplot(data = tips, aes(x = Party, y = Tip)) + 
  geom_jitter() + 
  labs(x = "Number of diners in party", 
       y = "Tips ($)",
       title = "Tips vs. Party")

p5 <- ggplot(data = tips, aes(x = Age, y = Tip)) + 
  geom_boxplot() + 
  labs(x = "Age of payer", 
       y = "Tips ($)",
       title = "Tips vs. Age")

p4 + p5
```

**We will use the number of diners in the party and age of the payer to understand variability in the tips.**

## Exercise 1

We fit a model of `Tip` using `Age`, `Party`, and the interaction between the two variables:

$$
\begin{aligned}
Tip_i = \beta_0 &+ \beta_1Party_i + \beta_2SenCit_i + \beta_3Yadult_i \\ 
&+ \beta_4Party_i \times SenCit_i + \beta_5 Party_i \times Yadult_i \\
&+ \epsilon_i
\end{aligned}
$$

a.  Which of the following is true for this model? Select all that apply.

    i.  The intercepts are the same for every level of `Age`.

    ii. The intercepts differ by `Age`.

    iii. The effect of `Party` is the same for every level of `Age`.

    iv. The effect of `Party` differs by `Age`.

b.  By how much does the intercept for tables with young adult payers differ from tables with middle age payers? Write the answer in terms of the $\beta$'s.

c.  Write the equation of the model for tables in which the payer is a senior citizen. Write the answer in terms of the $\beta$'s.

d.  Suppose you wish to test the hypotheses: $H_0: \beta_5 = 0 \text{ vs. }H_a: \beta_5 \neq 0$ . State what is being tested in terms of the effect of `Party`.

## Exercise 2

Use the `lm()` function to fit the model that includes `Age` and `Party` and the interaction between the two variables. Display the 90% confidence interval for the coefficients.

a.  State what the standard error for the coefficient for the main effect of `Party` means in the context of the data.

b.  Write code to show how the 90% confidence interval for the main effect of `Party` was computed.

c.  Based on the confidence interval, is there evidence of an association between the number of diners and the tip among tables with middle-aged payers? Explain.

d.  Explain what we know about the p-value for the coefficient of `Party` based on results from part (c).

## Exercise 3

Use the model fit in Exercise 2. Interpret the coefficient for `AgeSenCit:Party` in the context of the data.

## Exercise 4

Use the model fit in Exercise 2.

a.  Compute $SST$ and explain its meaning in the context of the data.

b.  Compute $SSR$ and explain its meaning in the context of the data.

c.  Compute $SSM$ and explain its meaning in the context of the data.

# Submission

You will submit the PDF documents for labs, homework, and exams in Gradescope as part of your final submission.

::: callout-warning
Before you wrap up the assignment, make sure all documents are updated on your GitHub repo. We will be checking these to make sure you have been practicing how to commit and push changes.

Remember -- you must turn in a PDF file to Gradescope by the end of your lab session.

If you write your responses to conceptual exercises by hand, you will need to combine your written work to the completed PDF for the applied exercises before submitting on Gradescope.

Instructions to combine PDFs:

- Preview (Mac): [support.apple.com/guide/preview/combine-pdfs-prvw43696/mac](https://support.apple.com/guide/preview/combine-pdfs-prvw43696/mac)

- Adobe (Mac or PC): [helpx.adobe.com/acrobat/using/merging-files-single-pdf.html](https://helpx.adobe.com/acrobat/using/merging-files-single-pdf.html)

- Get free access to Adobe Acrobat as a Duke student: [oit.duke.edu/help/articles/kb0030141/](https://oit.duke.edu/help/articles/kb0030141/)
:::

**To submit your assignment:**

- Access Gradescope through the menu on the [STA 221 Canvas site.](https://canvas.duke.edu/courses/84458)

- Click on the assignment, and you’ll be prompted to submit it.

- Mark the pages associated with each exercise. Each question in Gradescope should have at least one page associated with it.

- Select the first page of your .PDF submission to be associated with "Attendance & submitting lab" and "Workflow & formatting"

# Grading (30 pts)

This lab will be graded based on completion and workflow & formatting. The point breakdown is as follows:

| Component                   | Points             |
|-----------------------------|--------------------|
| Attendance & submitting lab | 10                 |
| Two exercises               | 16 (8 points each) |
| Workflow & formatting       | 4                  |

------------------------------------------------------------------------

We will grade two randomly selected exercises.

The “Workflow & formatting” grade is to assess the reproducible workflow and document format for the applied exercises. This includes having at least 3 informative commit messages, a neatly organized document with readable code, along with your name and the date updated in the YAML.
