---
title: "Lab 04: Exam 01 Review"
date: "February 15, 2026"
html-math-method:
  method: mathjax
  url: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-mml-chtml.js"
editor: visual
execute:
  warning: false
  message: false
---

::: callout-important
# Due date

This lab is due on **Sunday, February 15 at 11:59pm.** To be considered on time, the following must be done by the due date:

-   Final `.qmd` and `.pdf` files pushed to your team's GitHub repo
-   Final `.pdf` file submitted on Gradescope
:::

# Getting Started

-   Go to the [sta221-sp26](https://www.github.com/sta221-sp26) organization on GitHub. Click on the repo with the prefix **lab-04**. It contains the starter documents you need to complete the lab.

-   Clone the repo and start a new project in RStudio. See the [Lab 00 instructions](../labs/lab-00.html) for details on cloning a repo and starting a new project in R.

-   Each person on the team should clone the repository and open a new project in RStudio. Throughout the lab, each person should get a chance to make commits and push to the repo

# Packages

You will need the following packages for the lab:

```{r}
#| label: packages

library(tidyverse)
library(tidymodels) 
library(knitr)
library(patchwork)

# add other packages as needed
```

# Restaurant tips

**What factors are associated with the amount customers tip at a restaurant?** To answer this question, we will use data collected in 2011 by a student at St. Olaf who worked at a local restaurant.[^1]

[^1]: Dahlquist, Samantha, and Jin Dong. 2011. "The Effects of Credit Cards on Tipping." Project for Statistics 212-Statistics for the Sciences, St. Olaf College.

The variables we'll focus on for this analysis are

-   `Tip`: amount of the tip
-   `Party`: number of people in the party
-   `Age`: Age of the payer

View the data set to see the remaining variables.

```{r}
#| message: false

tips <- read_csv("data/tip-data.csv")
```

## Exploratory data analysis

```{r}
#| label: univariate-eda
#| warning: false

p1 <- ggplot(data = tips, aes(x = Tip)) + 
  geom_histogram(color = "white", binwidth = 2) +
  labs(x = "Tips ($)",
       title = "Tips at local restaurant")

p2 <- ggplot(data = tips, aes(x = Party)) + 
  geom_histogram(color = "white") +
  labs(x = "Party",
       title = "Number of diners in party") +
  xlim(c(0, 7))

p3 <- ggplot(data = tips, aes(x = Age)) + 
  geom_bar(color = "white") +
  labs(x = "",
       title = "Age of Payer") 

p1 / (p2 + p3)
```

```{r}
#| label: bivariate-eda

p4 <- ggplot(data = tips, aes(x = Party, y = Tip)) + 
  geom_jitter() + 
  labs(x = "Number of diners in party", 
       y = "Tips ($)",
       title = "Tips vs. Party")

p5 <- ggplot(data = tips, aes(x = Age, y = Tip)) + 
  geom_boxplot() + 
  labs(x = "Age of payer", 
       y = "Tips ($)",
       title = "Tips vs. Age")

p4 + p5
```

**We will use the number of diners in the party and age of the payer to understand variability in the tips.**

## Exercise 1

We will start with the main effects model that includes `Age` and `Party`.

a.  How many indicator variables for `Age` can we create from the data?

b.  How many indicator variables for `Age` will be in the regression model?

c.  Are the responses to parts a and b equal? If not, explain why not.

d.  Which of the following is true for this model? Select all that apply.

    i.  The intercepts are the same for every level of `Age`.
    ii. The intercepts differ by `Age`.
    iii. The effect of `Party` is the same for every level of `Age`.
    iv. The effect of `Party` differs by `Age`.

## Exercise 2

Consider the main effects model that includes `Age` and `Party`.

a.  What is the dimension of the design matrix $\mathbf{X}$ for the main effects model?

b.  Calculate the coefficient estimates $\hat{\boldsymbol{\beta}}$ directly from the data.

c.  Write the equation of the estimated regression model.

## Exercise 3

Consider the main effects model that includes `Age` and `Party`. Get $\mathbf{y}$ and $\mathbf{X}$ from the data.

a.  Use $\mathbf{y}$ and $\mathbf{X}$ to compute $\hat{\sigma}_{\epsilon}$ .

b.  Interpret $\hat{\sigma}_\epsilon$ in the context of the data.

c.  Compute $Var(\hat{\boldsymbol{\beta}})$.

d.  You wish to test whether there is a linear relationship between tips and the number of diners in the party, after adjusting for the age of the payer. Compute the test statistic. Explain what it means in the context of the data.

e.   Use the code below to compute the p-value and explain what it means in the context of the data.

f.  State the conclusion from the test in the context of the data.

```{r}
#| eval: false
2 * pt(abs(test-statistic), df, lower.tail = FALSE)
```

## Exercise 4

Consider the main effects model that includes `Age` and `Party`. Get $\mathbf{y}$ and $\mathbf{X}$ from the data.

a.  Use $\mathbf{y}$ and $\mathbf{X}$ to compute $R^2$. Interpret this value in the context of the data.

b.  Use $\mathbf{y}$ and $\mathbf{X}$ to compute $RMSE$. Interpret this value in the context of the data.

## Exercise 5

You decide to add an interaction effect between `Age` and `Party` to the model and fit a model of the following form:

$$
\begin{aligned}
Tip_i = \beta_0 &+ \beta_1Party_i + \beta_2SenCit_i + \beta_3Yadult_i \\ 
&+ \beta_4Party_i \times SenCit_i + \beta_5 Party_i \times Yadult_i \\
&+ \epsilon_i
\end{aligned}
$$

a.  Which of the following is true for this model? Select all that apply.

    i.  The intercepts are the same for every level of `Age`.

    ii. The intercepts differ by `Age`.

    iii. The effect of `Party` is the same for every level of `Age`.

    iv. The effect of `Party` differs by `Age`.

b.  By how much does the intercept for tables with young adult payers differ from tables with middle age payers? Write the answer in terms of the $\beta$'s.

c.  Write the equation of the model for tables in which the payer is a senior citizen.

d.  Suppose you wish to test the hypotheses: $H_0: \beta_5 = 0 \text{ vs. }H_a: \beta_5 \neq 0$ . State what is being tested in terms of the effect of `Party`.

## Exercise 6

Use the `lm()` function to fit the model that includes `Age` and `Party` and the interaction between the two variables. Display the 90% confidence interval for the coefficients.

a.  The standard error for `AgeSenCit` is 0.784. State what this value means in the context of the data.

b.  Write code to show how the 90% confidence interval for `AgeSenCit` was computed.

c.  Based on the confidence interval, is there evidence that tables with senior citizen payers tip differently on average than tables with middle age payers?

# Submission

You will submit the PDF documents for labs, homework, and exams in to Gradescope as part of your final submission.

::: callout-warning
Before you wrap up the assignment, make sure all documents are updated on your GitHub repo. We will be checking these to make sure you have been practicing how to commit and push changes.

Remember – you must turn in a PDF file to the Gradescope page before the submission deadline for full credit.
:::

To submit your assignment:

-   Access Gradescope through the menu on the [STA 221 Canvas site.](https://canvas.duke.edu/courses/73984)

-   Click on the assignment, and you’ll be prompted to submit it.

-   Select the name of **every** team member.

-   Mark the pages for the *"Exercises"* question.

-   Select the first page of your .PDF submission to be associated with the *"Team agreement"* and *“Workflow & formatting”* questions.

# Grading (10 pts)

This lab will be graded based on completion and workflow & formatting. The point breakdown is as follows:

| Component             | Points |
|-----------------------|--------|
| Complete exercises    | 8      |
| Workflow & formatting | 2      |

: **Completion:** Points will be awarded for completion based on the following:

-   8 pts: Complete all exercises

-   7 pts: Complete 5 exercises

-   6 pts: Complete 4 exercises

-   5 pts: Complete 3 exercises

-   4 pts: Complete 2 exercises

-   0 pts: Complete \< 2 exercises

**Workflow & formatting**

The “Workflow & formatting” grade is to assess the reproducible workflow and collaboration. This includes having at least one meaningful commit from each team member and updating the team name and date in the YAML.

-   2 pts: Meet all criteria

-   1 pt: Meet some criteria

-   0 pt: Meet no criteria
